Weighted Model Counting with Conditional Weights for Bayesian Networks

### 2020-11-25 paper2/paper.tex
Transfer the paper to the new format

\def\year{2021}\relax
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai21}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{graphicx}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
\pdfinfo{
/Title (Weighted Model Counting with Conditional Weights for Bayesian Networks)
/Author ()
/TemplateVersion (2021.1)
} %Leave this
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case.
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands,
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.

% The file aaai21.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

\usepackage{amsthm}
\usepackage{thmtools,thm-restate}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage{complexity}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[binary-units]{siunitx}
\usepackage[inline]{enumitem}
\usetikzlibrary{cd}
\usetikzlibrary{bayesnet}
\usetikzlibrary{calc}
\DeclareMathOperator{\im}{im}
\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\makeatletter
\newcommand{\nosemic}{\renewcommand{\@endalgocfline}{\relax}}% Drop semi-colon ;
\newcommand{\dosemic}{\renewcommand{\@endalgocfline}{\algocf@endline}}% Reinstate semi-colon ;
\newcommand{\pushline}{\Indp}% Indent
\newcommand{\popline}{\Indm\dosemic}% Undent
\makeatother

\title{Weighted Model Counting with Conditional Weights for Bayesian Networks}
%\author{
%  Paulius Dilkas, \textsuperscript{\rm 1}
%  Vaishak Belle \textsuperscript{\rm 1,2} \\
%}
%\affiliations{
%    \textsuperscript{\rm 1} University of Edinburgh \\
%    \textsuperscript{\rm 2} Alan Turing Institute \\
%    p.dilkas@sms.ed.ac.uk, vaishak@ed.ac.uk
%}
\begin{document}

@@

% TODO: Feedback: our empirical evaluations show a x% improvement

@@

\paragraph{WMC for Bayesian networks.} Hitherto, four techniques have been
proposed for encoding Bayesian networks into instances of WMC. We will identify
them based on the initials of authors as well as publications years:
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05},
and \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}. Below we summarise the
observed performance differences among them. \citet{DBLP:conf/aaai/SangBK05}
claim that \texttt{sbk05} is a smaller encoding than \texttt{d02} with respect
to both the number of clauses and the number of variables but provide no
experimental comparison. \citet{DBLP:conf/ijcai/ChaviraD05} compare
\texttt{cd05} with \texttt{d02} by measuring the time it takes to compile either
encoding into an arithmetic circuit (but do not measure inference time). The
more recent encoding \texttt{cd05} always compiles faster and results in a
smaller arithmetic circuit (as measured by the number of edges). In their
subsequent paper, the same authors perform two sets of experiments (that are
relevant to this summary) \cite{DBLP:conf/sat/ChaviraD06}. First, they compile
\texttt{cd05} and \texttt{cd06} encodings into d-DNNF (i.e., deterministic
decomposable negation normal form \cite{DBLP:journals/jancl/Darwiche01}),
measuring both compilation time and numbers of edges in the d-DNNF diagram. The
results are mostly in favour of \texttt{cd06}. Second, they compare the
inference time of \texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04}
with the compile times of \texttt{cd05} and \texttt{cd06}, but only on five
(types of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed. Sometimes
\texttt{cd06} is orders of magnitude faster than \texttt{sbk05}, sometimes
slightly slower. The performance difference between \texttt{sbk05} and
\texttt{cd05} is even harder to judge: \texttt{sbk05} is better on three out of
five instances and worse on the remaining two. Based on this description, one
would expect \texttt{cd06} to be faster than both \texttt{cd05} and
\texttt{sbk05}, both of which should be faster than \texttt{d02}. The
experiments in \cref{sec:experiments}, however, strongly disagree with this
prediction, showing that the quality of an encoding depends strongly on the
underlying search algorithm or compilation technique.

@@

\begin{table*}
  \caption{Notation for a logic with two atoms. The elements in both columns are
    listed in the same order.}
  \label{tbl:notation_example}
  \centering
  \begin{tabular}{lcc}
    \toprule
    Name in logic & Boolean-algebraic notation & Set-theoretic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\emptyset, \{a\}, \{b\}, \{a, b\}$ \\
    & $\top$ & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg a \lor \neg b, a \to b$ & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ \\
    & $b \to a, a \lor b$ & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg b, \neg a, a \leftrightarrow b$ & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ \\
    & $(a \land \neg b) \lor (b \land \neg a), a, b$ & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ \\
    & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\bot$ & $\emptyset$ \\
    \bottomrule
  \end{tabular}
\end{table*}

@@

We also consider the space of all functions from any Boolean algebra to
$\mathbb{R}_{\ge 0}$ together with some operations on those functions. They will
be instrumental in defining WMC as a measure in \cref{sec:wmc_as_measure} and
can be efficiently represented using ADDs. Furthermore, all of the operations
are supported by the CUDD \cite{somenzi1998cudd} package that is used by ADDMC
for ADD manipulation \cite{DBLP:conf/aaai/DudekPV20}. The definitions of
multiplication and projection are as defined by
\citet{DBLP:conf/aaai/DudekPV20}, while others are new.

@@

\begin{algorithm}[t]
  \caption{Encoding a Bayesian network}
  \label{alg:encoding}
  \KwData{vertices $\mathcal{V}$, probability distribution $\Pr$}
  \KwResult{$\phi\colon 2^U \to \mathbb{R}_{\ge 0}$}
  $\phi \gets 1$\;
  \For{$X \in \mathcal{V}$}{
    \textit{let} $\mathrm{pa}(X) = \{ Y_1, \dots, Y_n \}$\;
    $\mathrm{CPT}_X \gets 0$\;
    \eIf{$|\im X| = 2$}{
      \For{$(y_i)_{i=1}^n \in \prod_{i = 1}^n \im Y_i$}{
        $p_1 \gets \Pr(X = 1 \mid y_1, \dots, y_n)$\;
        $p_0 \gets \Pr(X \ne 1 \mid y_1, \dots, y_n)$\;
        \nosemic $\mathrm{CPT}_X \gets \mathrm{CPT}_X$\;
        \pushline $+ p_1[\lambda_{X=1}] \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
        \dosemic $+ p_0 \overline{[\lambda_{X=1}]} \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
      }
    }{
      \textit{let} $\im X = \{ x_1, \dots, x_m \}$\;
      \For{$x \in \im X$ {\rm \textbf{and}} $(y_i)_{i=1}^n \in \prod_{i = 1}^n
        \im Y_i$}{
        $p_x \gets \Pr(X = x \mid y_1, \dots, y_n)$\;
        \nosemic $\mathrm{CPT}_X \gets \mathrm{CPT}_X$\;
        \pushline $+ p_x[\lambda_{X=x}] \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
        \dosemic $+ \overline{[\lambda_{X=x}]} \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
      }
      \nosemic $\mathrm{CPT}_X \gets \mathrm{CPT}_X \cdot \left( \sum_{i=1}^m
        [\lambda_{X = x_i}] \right)$\;
      \pushline\dosemic $\cdot \prod_{i=1}^m \prod_{j=i+1}^m
      (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})$\;
    }
    $\phi \gets \phi \cdot \mathrm{CPT}_X$\;
  }
  \Return{$\phi$}\;
\end{algorithm}

@@

For experimental data, we use the Bayesian networks available with Ace and
Cachet\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}, most of
which happen to be binary. We classify them into the following seven categories:
\begin{itemize*}
\item DQMR and
\item Grid networks as described by \citet{DBLP:conf/aaai/SangBK05},
\item Mastermind, and
\item Random Blocks from the work of \citet{DBLP:journals/ijar/ChaviraDJ06},
\item remaining binary Bayesian networks that include Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f}, and
\item non-binary classic Bayesian networks (\texttt{alarm}, \texttt{diabetes},
  \texttt{hailfinder}, \texttt{mildew}, \texttt{munin1}--\texttt{4},
  \texttt{pathfinder}, \texttt{pigs}, \texttt{water}).
\end{itemize*}
We run ADDMC with each of the five encodings once on each Bayesian network.

\begin{figure}
  \centering
  \input{cumulative.tex}%
  \caption{Cumulative number of instances solved by ADDMC over time using each
    encoding}
  \label{fig:cumulative}
\end{figure}

% TODO: (needed to follow formatting requirements)
% 1. Remove the \input command.
% 2. R should compile the tex files to be standalone.
% 3. Edit the tex files to replace RGB colours with CMYK.
% 4. Compile the tex files to PDFs.
% 5. Include the PDFs (look up how to do that in the formatting guidelines).
% 6. make sure that there are no duplicates in the bib files (look at LaTex output)

\begin{table}
  \centering
  \caption{The numbers of instances (out of 1466) solved by ADDMC with each
    encoding (uniquely, faster than with others, and in total) under the
    described constraints}%
  \begin{tabular}{lrrr}
    \toprule
    Encoding & Unique & Fastest & Total \\
    \midrule
    \texttt{cd05} & 2 & 3 & 380 \\
    \texttt{cd06} & 0 & 0 & 363 \\
    \texttt{cw} & 234 & 935 & 980 \\
    \texttt{d02} & 40 & 111 & 797 \\
    \texttt{sbk05} & 12 & 44 & 765 \\
    \bottomrule
  \end{tabular}
  \label{tbl:tallies}
\end{table}

\begin{figure*}
  \centering
  \input{scatter.tex}%
  \caption{ADDMC inference time using \texttt{cw} compared to \texttt{d02}
    (left) and \texttt{sbk05} (right) for each instance in all data sets.}
  \label{fig:scatter}
\end{figure*}

@@

\begin{table}
  \centering
  \caption{Asymptotic upper bounds on the numbers of variables and clauses/ADDs
    for each encoding}
  \label{tbl:asymptotes}
  \begin{tabular}{lcc}
    \toprule
    Encoding(s) & Variables & Clauses/ADDs \\
    \midrule
    \texttt{cd05}, \texttt{cd06}, \texttt{sbk05} & $\mathcal{O}(nv^{d+1})$ & $\mathcal{O}(nv^{d+1})$ \\
    \texttt{cw} & $\mathcal{O}(nv)$ & $\mathcal{O}(nv^2)$ \\
    \texttt{d02} & $\mathcal{O}(nv^{d+1})$ & $\mathcal{O}(ndv^{d+1})$ \\
    \bottomrule
  \end{tabular}
\end{table}

@@

While we chose to use ADDMC \cite{DBLP:conf/aaai/DudekPV20} as the WMC algorithm
and Bayesian networks as a canonical example of a probabilistic inference task,
these are only examples meant to illustrate the broader idea that choosing a
more expressive representation of weights can outperform increasing the size of
the problem to keep the weights simple. Indeed, in this work, we have provided a
new theoretical perspective on the expressive power of WMC and illustrated the
empirical benefits of that perspective. The same idea can be adapted to
search-based solvers such as Cachet \cite{DBLP:conf/sat/SangBBKP04} as well as
to other inference problem domains such as probabilistic programs
\cite{DBLP:journals/tplp/FierensBRSGTJR15,DBLP:journals/corr/abs-2005-09089}.

@@

\newpage
\bibliography{paper}

\newpage
\appendix
\section{Proofs}

@@

\correctness*
\begin{proof}
  Let $\mathcal{V} = \{ X, Y_1, \dots, Y_n \}$. Then
  \begin{align*}
    (\exists_U (\phi \cdot [\lambda_{X=x}]))(\emptyset) &= \sum_{T \in 2^U} (\phi \cdot [\lambda_{X=x}])(T) \\
                                                        &= \sum_{\lambda_{X=x} \in T \in 2^U} \phi(T) \\
                                                        &= \sum_{\lambda_{X=x} \in T \in 2^U} \left( \prod_{Y \in \mathcal{V}} \mathrm{CPT}_Y \right)(T) \\
                                                        &= \sum_{(y_i)_{i=1}^n \in \prod_{i=1}^n \im Y_i} \Pr(x, y_1, \dots, y_n) \\
                                                        &= \Pr(X = x)
  \end{align*}
  by:
  \begin{itemize}
  \item the proof of Theorem~1 by \citet{DBLP:conf/aaai/DudekPV20};
  \item if $\lambda_{X=x} \not\in T \in 2^U$, then $(\phi \cdot
    [\lambda_{X=x}])(T) = \phi(T) \cdot [\lambda_{X=x}](T \cap \{
    \lambda_{X=x} \}) = \phi(T) \cdot 0 = 0$;
  \item \cref{lemma:full_distribution};
  \item marginalisation of a probability distribution.
  \end{itemize}
\end{proof}

>>>>

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
\usepackage{ijcai21}
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\urlstyle{same}

\usepackage{complexity}
\usepackage[capitalise]{cleveref}
\usepackage{amsfonts}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[binary-units]{siunitx}
\usepackage[inline]{enumitem}
\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{tikz}

% the following package is optional:
%\usepackage{latexsym}

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}

\DeclareMathOperator{\im}{im}
\newcommand{\etal}{\textit{et al}.}
\Crefname{algocf}{Algorithm}{Algorithms}
\Crefname{figure}{Figure}{Figures}

\makeatletter
\newcommand{\nosemic}{\renewcommand{\@endalgocfline}{\relax}}% Drop semi-colon ;
\newcommand{\dosemic}{\renewcommand{\@endalgocfline}{\algocf@endline}}% Reinstate semi-colon ;
\newcommand{\pushline}{\Indp}% Indent
\newcommand{\popline}{\Indm\dosemic}% Undent
\makeatother

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

%PDF Info Is REQUIRED.
\pdfinfo{
/TemplateVersion (IJCAI.2021.0)
}

\title{Weighted Model Counting with Conditional Weights for Bayesian Networks}

% Single author syntax
% \author{
%     Zhi-Hua Zhou
%     \affiliations
%     Nanjing University
%     \emails
%     pcchair@ijcai-21.org
% }

% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
% Check the ijcai21-multiauthor.tex file for detailed instructions
% \iffalse
% \author{
% First Author$^1$
% \and
% Second Author$^2$\and
% Third Author$^{2,3}$\And
% Fourth Author$^4$
% \affiliations
% $^1$First Affiliation\\
% $^2$Second Affiliation\\
% $^3$Third Affiliation\\
% $^4$Fourth Affiliation
% \emails
% \{first, second\}@example.com,
% third@other.example.com,
% fourth@example.com
% }
% \fi

\begin{document}

@@

@@

\paragraph{WMC for Bayesian networks.} Hitherto, four techniques have been
proposed for encoding Bayesian networks into instances of WMC. We will identify
them based on the initials of authors as well as publications years:
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05},
and \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}. Below we summarise the
observed performance differences among them. Sang
\etal{}~\shortcite{DBLP:conf/aaai/SangBK05} claim that \texttt{sbk05} is a
smaller encoding than \texttt{d02} with respect to both the number of clauses
and the number of variables but provide no experimental comparison. Chavira and
Darwiche~\shortcite{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with
\texttt{d02} by measuring the time it takes to compile either encoding into an
arithmetic circuit (but do not measure inference time). The more recent encoding
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\cite{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \cite{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04} with the compile
times of \texttt{cd05} and \texttt{cd06}, but only on five (types of) instances.
In these experiments, \texttt{cd06} is always faster than \texttt{cd05}, while
the comparison with \texttt{sbk05} is mixed. Sometimes \texttt{cd06} is orders
of magnitude faster than \texttt{sbk05}, sometimes slightly slower. The
performance difference between \texttt{sbk05} and \texttt{cd05} is even harder
to judge: \texttt{sbk05} is better on three out of five instances and worse on
the remaining two. Based on this description, one would expect \texttt{cd06} to
be faster than both \texttt{cd05} and \texttt{sbk05}, both of which should be
faster than \texttt{d02}. The experiments in \cref{sec:experiments}, however,
strongly disagree with this prediction, showing that the quality of an encoding
depends strongly on the underlying search algorithm or compilation technique.

@@

\begin{table*}
  \centering
  \begin{tabular}{lcc}
    \toprule
    Name in logic & Boolean-algebraic notation & Set-theoretic notation \\
    \midrule
    Atoms (elements of $U$) & $a, b$ & $a, b$ \\
    \rowcolor{gray!10} Models (elements of $2^U$) & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\emptyset, \{a\}, \{b\}, \{a, b\}$ \\
    & $\top$ & $\{ \emptyset, \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg a \lor \neg b, a \to b$ & $\{ \emptyset, \{a\}, \{b\} \}, \{ \emptyset, \{a\}, \{a, b\} \}$ \\
    & $b \to a, a \lor b$ & $\{ \emptyset, \{b\}, \{a, b\} \}, \{ \{a\}, \{b\}, \{a, b\} \}$ \\
    & $\neg b, \neg a, a \leftrightarrow b$ & $\{\emptyset, \{a\}\}, \{\emptyset, \{b\}\}, \{\emptyset, \{a, b\}\}$ \\
    & $(a \land \neg b) \lor (b \land \neg a), a, b$ & $\{\{a\}, \{b\}\}, \{\{a\}, \{a, b\}\}, \{\{b\}, \{a, b\}\}$ \\
    & $\neg a \land \neg b, a \land \neg b, \neg a \land b, a \land b$ & $\{\emptyset\}, \{\{a\}\}, \{\{b\}\}, \{\{a, b\}\}$ \\
    \multirow{-7}{*}{Formulas (elements of $2^{2^U}$)} & $\bot$ & $\emptyset$ \\
    \bottomrule
  \end{tabular}
  \caption{Notation for a logic with two atoms. The elements in both columns are
    listed in the same order.}
  \label{tbl:notation_example}
\end{table*}

@@

We also consider the space of all functions from any Boolean algebra to
$\mathbb{R}_{\ge 0}$ together with some operations on those functions. They will
be instrumental in defining WMC as a measure in \cref{sec:wmc_as_measure} and
can be efficiently represented using ADDs. Furthermore, all of the operations
are supported by the CUDD \cite{somenzi1998cudd} package that is used by ADDMC
for ADD manipulation \cite{DBLP:conf/aaai/DudekPV20}. The definitions of
multiplication and projection are as defined by
Dudek \etal{}~\shortcite{DBLP:conf/aaai/DudekPV20}, while others are new.

@@

\begin{algorithm}[t]
  \KwData{vertices $\mathcal{V}$, probability distribution $\Pr$}
  \KwResult{$\phi\colon 2^U \to \mathbb{R}_{\ge 0}$}
  $\phi \gets 1$\;
  \For{$X \in \mathcal{V}$}{
    \textit{let} $\mathrm{pa}(X) = \{ Y_1, \dots, Y_n \}$\;
    $\mathrm{CPT}_X \gets 0$\;
    \eIf{$|\im X| = 2$}{
      \For{$(y_i)_{i=1}^n \in \prod_{i = 1}^n \im Y_i$}{
        $p_1 \gets \Pr(X = 1 \mid y_1, \dots, y_n)$\;
        $p_0 \gets \Pr(X \ne 1 \mid y_1, \dots, y_n)$\;
        \nosemic $\mathrm{CPT}_X \gets \mathrm{CPT}_X$\;
        \pushline $+ p_1[\lambda_{X=1}] \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
        \dosemic $+ p_0 \overline{[\lambda_{X=1}]} \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
      }
    }{
      \textit{let} $\im X = \{ x_1, \dots, x_m \}$\;
      \For{$x \in \im X$ {\rm \textbf{and}} $(y_i)_{i=1}^n \in \prod_{i = 1}^n
        \im Y_i$}{
        $p_x \gets \Pr(X = x \mid y_1, \dots, y_n)$\;
        \nosemic $\mathrm{CPT}_X \gets \mathrm{CPT}_X$\;
        \pushline $+ p_x[\lambda_{X=x}] \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
        \dosemic $+ \overline{[\lambda_{X=x}]} \cdot \prod_{i=1}^n
        [\lambda_{Y_i=y_i}]$\;
      }
      \nosemic $\mathrm{CPT}_X \gets \mathrm{CPT}_X \cdot \left( \sum_{i=1}^m
        [\lambda_{X = x_i}] \right)$\;
      \pushline\dosemic $\cdot \prod_{i=1}^m \prod_{j=i+1}^m
      (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})$\;
    }
    $\phi \gets \phi \cdot \mathrm{CPT}_X$\;
  }
  \Return{$\phi$}\;
  \caption{Encoding a Bayesian network}
  \label{alg:encoding}
\end{algorithm}

@@

For experimental data, we use the Bayesian networks available with Ace and
Cachet\footnote{\url{https://www.cs.rochester.edu/u/kautz/Cachet/}}, most of
which happen to be binary. We classify them into the following seven categories:
\begin{itemize*}
\item DQMR and
\item Grid networks as described by Sang
  \etal{}~\shortcite{DBLP:conf/aaai/SangBK05},
\item Mastermind, and
\item Random Blocks from the work of Chavira
  \etal{}~\shortcite{DBLP:journals/ijar/ChaviraDJ06},
\item remaining binary Bayesian networks that include Plan Recognition
  \cite{DBLP:conf/aaai/SangBK05}, Friends and Smokers, Students and Professors
  \cite{DBLP:journals/ijar/ChaviraDJ06}, and \texttt{tcc4f}, and
\item non-binary classic Bayesian networks (\texttt{alarm}, \texttt{diabetes},
  \texttt{hailfinder}, \texttt{mildew}, \texttt{munin1}--\texttt{4},
  \texttt{pathfinder}, \texttt{pigs}, \texttt{water}).
\end{itemize*}
We run ADDMC with each of the five encodings once on each Bayesian network.

\begin{figure} % TODO: update
  \centering
  \input{../paper1/cumulative.tex}%
  \caption{Cumulative number of instances solved by ADDMC over time using each
    encoding}
  \label{fig:cumulative}
\end{figure}

\begin{table}
  \centering
  \begin{tabular}{lrrr}
    \toprule
    Encoding & Unique & Fastest & Total \\
    \midrule
    \texttt{cd05} & 2 & 3 & 380 \\
    \texttt{cd06} & 0 & 0 & 363 \\
    \texttt{cw} & 234 & 935 & 980 \\
    \texttt{d02} & 40 & 111 & 797 \\
    \texttt{sbk05} & 12 & 44 & 765 \\
    \bottomrule
  \end{tabular}
  \caption{The numbers of instances (out of 1466) solved by ADDMC with each
    encoding (uniquely, faster than with others, and in total) under the
    described constraints}%
  \label{tbl:tallies}
\end{table}

\begin{figure*} % TODO: update
  \centering
  \input{../paper1/scatter.tex}%
  \caption{ADDMC inference time using \texttt{cw} compared to \texttt{d02}
    (left) and \texttt{sbk05} (right) for each instance in all data sets.}
  \label{fig:scatter}
\end{figure*}

@@

\begin{table}
  \centering
  \begin{tabular}{lcc}
    \toprule
    Encoding(s) & Variables & Clauses/ADDs \\
    \midrule
    \texttt{cd05}, \texttt{cd06}, \texttt{sbk05} & $\mathcal{O}(nv^{d+1})$ & $\mathcal{O}(nv^{d+1})$ \\
    \texttt{cw} & $\mathcal{O}(nv)$ & $\mathcal{O}(nv^2)$ \\
    \texttt{d02} & $\mathcal{O}(nv^{d+1})$ & $\mathcal{O}(ndv^{d+1})$ \\
    \bottomrule
  \end{tabular}
  \caption{Asymptotic upper bounds on the numbers of variables and clauses/ADDs
    for each encoding}
  \label{tbl:asymptotes}
\end{table}

@@

While we chose to use ADDMC \cite{DBLP:conf/aaai/DudekPV20} as the WMC algorithm
and Bayesian networks as a canonical example of a probabilistic inference task,
these are only examples meant to illustrate the broader idea that choosing a
more expressive representation of weights can outperform increasing the size of
the problem to keep the weights simple. Indeed, in this work, we have provided a
new theoretical perspective on the expressive power of WMC and illustrated the
empirical benefits of that perspective. The same idea can be adapted to other
inference problem domains such as probabilistic programs
\cite{DBLP:journals/tplp/FierensBRSGTJR15,DBLP:journals/corr/abs-2005-09089} as
well as to search-based solvers such as Cachet \cite{DBLP:conf/sat/SangBBKP04},
although the extension may be less straightforward for other WMC techniques.

@@

\appendix
\section{Proofs}

@@

\correctness*
\begin{proof}
  Let $\mathcal{V} = \{ X, Y_1, \dots, Y_n \}$. Then
  \begin{align*}
    (\exists_U (\phi \cdot [\lambda_{X=x}]))(\emptyset) &= \sum_{T \in 2^U} (\phi \cdot [\lambda_{X=x}])(T) \\
                                                        &= \sum_{\lambda_{X=x} \in T \in 2^U} \phi(T) \\
                                                        &= \sum_{\lambda_{X=x} \in T \in 2^U} \left( \prod_{Y \in \mathcal{V}} \mathrm{CPT}_Y \right)(T) \\
                                                        &= \sum_{(y_i)_{i=1}^n \in \prod_{i=1}^n \im Y_i} \Pr(x, y_1, \dots, y_n) \\
                                                        &= \Pr(X = x)
  \end{align*}
  by:
  \begin{itemize}
  \item the proof of Theorem~1 by Dudek
    \etal{}~\shortcite{DBLP:conf/aaai/DudekPV20};
  \item if $\lambda_{X=x} \not\in T \in 2^U$, then $(\phi \cdot
    [\lambda_{X=x}])(T) = \phi(T) \cdot [\lambda_{X=x}](T \cap \{
    \lambda_{X=x} \}) = \phi(T) \cdot 0 = 0$;
  \item \cref{lemma:full_distribution};
  \item marginalisation of a probability distribution.
  \end{itemize}
\end{proof}

\bibliographystyle{named}
\bibliography{paper}
